NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PE and PG lipids

### 2019-10-07 Manuscript/manuscriptPGPE.tex
Update PCvsPG results
Update text and figures

\subsection{Lipid headgroup interactions in PC:PE and PC:PG mixtures}
According to the electrometer concept, the headgroup order parameters increase
with the addition of negatively charged PG lipids, but
%decrease when mixed with positively charged surfactants and
are not affected by the addition of zwitterionic PE lipids (Fig.~\ref{HGorderparametersPCvsPEPG}) \cite{seelig87, scherer87}. % or cholesterol.
%In addition to the results summarized in Fig. \ref{HGorderparametersPCvsPEPSPGchol},
%also mixtures of PC with negatively charged PI, CL, PA, and zwitterionic SM
%follow the eletrometer concept \cite{scherer87}.
%to mixtures of differently charged
%lipids are collected from different experiments in Fig. \ref{HGorderparametersPCvsPEPSPGchol}.
%%\section{PC lipid headgroup response to different mixtures in experiments}
%As shown in Fig. \ref{HGorderparametersPCvsPEPSPGchol}, order parameters of PC
%headgroup behave in various lipid mixtures as expected from the electrometer concept \cite{seelig87, scherer87},
%i.e., order parameters increase when anionic lipids are mixed with PC and decrease with cationic
%surfactants. The changes with the addition of neutral lipids is significantly smaller.
\begin{figure*}[]
  \centering
  \includegraphics[width=16.0cm]{../Figs/HGorderparametersPCvsPEPG.eps}
  \caption{\label{HGorderparametersPCvsPEPG}
    Modulation of PC headgroup order parameters with increasing amount of PE (left) and PG (right) in bilayer
    from experiments \cite{scherer87,macdonald87} and simulations with different force fields.
    Signs are determined as discussed in \cite{botan15,ollila16}.
  }
  \todo{Data for CHARMM, Lipid17, Slipids, and also GROMOS also without ions is now available.
    This data should be updated into this plot}
  %\todo{Simulation of CHARMM36 at 298K should be maybe rerun with Gromacs 5.}
\end{figure*}
%The headgroup order parameters for PC lipids (POPC and DOPC)
%mixed with PE, PS and PG lipids are shown in Fig. \ref{HGorderparametersPCvsPEPSPG}
%from different simulation model and experiments \cite{scherer87} with different
%mole fractions. As already discussed previosly, the PC lipid headgroup behaviour
%follows the electrometer concept in experiments when mixed with other lipids, i.e., the order
%parameters increase when mixed with negatively charged lipids (PS, PI, CL, PA and PG)
%remains almost unchaged when mixed with neutral lipids (PE and SM) \cite{scherer87}.
This is roughly observed also in simulations here, althouht the changes in PC headgroup with
PE are slightly overestimated and changes with PG lipid are slightly underestimated.
The underestimated increase of PC headgroup order parameters with the addition of
negaticely charged lipids was previously interepreted to indicate overestimated
counterion binding affinity, which overcompensates the effect of negative charge \cite{NMRlipidsIVps}.
Thus, the counterion binding to PG lipid bilayers (Fig. \ref{CIdensPG}) may be slightly
overestimated also in simulations here.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/CIdensPG.eps}
  \caption{\label{CIdensPG}
    Counterion densities and area per lipids from simulations with PG lipids.
    Experimental area for POPG at 303 K is 66.1 {\AA}$^2$ and 67 {\AA}$^2$ for DPPC at 323 K \cite{pan12b}.
  }
  \todo{Data for GROMOS is also available without ions, but I could not analyze it, probably because tpr file was not correct.}
\end{figure}

>>>>

\subsection{Lipid headgroup interactions in PC:PE and PC:PG mixtures and counterion binding}
\begin{figure*}[]
  \centering
  \includegraphics[width=16.0cm]{../Figs/HGorderparametersPCvsPEPG.eps}
  \caption{\label{HGorderparametersPCvsPEPG}
    Modulation of POPC headgroup order parameters with increasing amount of POPE (left) and POPG (right) in bilayer
    from experiments \cite{scherer87,macdonald87} and simulations with different force fields.
    Signs are determined as discussed in \cite{botan15,ollila16}.
  }
  \todo{Data for CHARMM without ions to be added once this issue is solved: https://github.com/NMRLipids/MATCH/issues/83}
  %\todo{Simulation of CHARMM36 at 298K should be maybe rerun with Gromacs 5.}
\end{figure*}
According to the electrometer concept, the headgroup order parameters increase
with the addition of negatively charged PG lipids, but
are not affected by the addition of zwitterionic PE and SM lipids
or cholesterol (Fig.~\ref{HGorderparametersPCvsPEPG}) \cite{seelig87, scherer87,antila18}. 
The different responses of PC headgroup order parameters to PE and PG are roughly
reproduced in CHARMM36 simulations, although the changes are slightly
overestimated upon addition of PE and underestimated upon addition of PG.
Interestingly, the response of POPC headgroup order parameters to POPE is
close to experiments also in Berger-OPLS simulations, even though the response
of Berger force field to cholesterol was significantly overestimated \cite{botan15}.
In all force fields except Slipids, the $\alpha$-carbon order parameters of different
hydrogens are responsing differently when mixed with PE or PG lipids \todo{Maybe we should figure out what is the reason for this}.
\todo{Maybe we should analyze the P-N vector angle from different simulations}.

In our previous study about PS lipids \cite{antila18}, we related the underestimated increase of PC
headgroup order parameters upon addition of negaticely charged lipids
%was previously interepreted
to the overestimated counterion binding affinity, which would overcompensates the effect of negative charge.
Similar correlation is observed here in other than GROMOS-CKP simulation:
Lipid 17 simulation exhibits the stronger counter-ion binding affinitiy to the pure POPG bilayer (Fig. \ref{CIdensPG})
and larger decrease in POPC $\beta$ carbon order parameter upon addition of POPG than other simulations.
This, together with the low area per molecule in Lipid17 (59.5 {\AA} vs. 66.1 {\AA} in experiments) suggests
that the counterion binding this model is too strong.
However, $\beta$-carbon order parameter of POPC in GROMOS-CKP simulations is in line with experiments and
increases more than in CHARMM36 and Slipids simulations even though the binding affinities in these
simulations are simialar. On the other hand, the different responses of different hydrogens in $\alpha$-carbons
complicate the interpretation. 
Thus, the counterion binding to PG lipid bilayers (Fig. \ref{CIdensPG}) may be slightly
overestimated also in simulations here.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/CIdensPG.eps}
  \caption{\label{CIdensPG}
    Counterion densities and area per lipids from simulations with PG lipids.
    Experimental area for POPG at 303 K is 66.1 {\AA}$^2$ and 67 {\AA}$^2$ for DPPC at 323 K \cite{pan12b}.
  }
  \todo{Data for GROMOS is also available without ions, but I could not analyze it, probably because tpr file was not correct.}
\end{figure}
